Modeling of Isobutane Alkylation with Mixed C4 Olefins and Sulfuric Acid as Catalyst Using the Asynchronous Global Optimization Algorithm

### 2021-12-12 2022/PAVT/PAVT2022.tex
UNN part translated

%Часть УГНТУ
Currently, there is a tendency to improve the environmental characteristics of automobile fuel while maintaining a high octane number. Sulfuric acid alkylation of isobutane with olefins makes it possible to obtain a high-octane component of gasoline with a minimum content of aromatic hydrocarbons. The alkylate, which is produced by the alkylation of isobutane with C3-C5 olefins in the presence of strong acid, has the advantages of high octane number, low vapor pressure, and zero content of olefins and aromatics that allow it to be a desirable blending component for high-quality gasoline. Alkylates will continue to act as a desirable blending component for high-quality gasoline as the quality of gasoline continues to increase [ссылка на китайскую статью]. Therefore, it is a significant process of the modern refinery [ссылка на catalyst]. To optimize the chemical process in industry, it is necessary to first develop its model, which in this case means building a mathematical model of the chemical process and then its kinetic model, that is, numerically calculate the kinetic constants of the reaction. 
%Стыковочный абзац
Как правило, найти кинетические константы реакций аналитическим путем не представляется возможным, поэтому возникает необходимость в разработке и использовании численных методов их поиска (см., например, [ссылки (Наша совместная статья Суперкомп дни 2021 ??)]). В этом случае критерий качества найденного решения (целевая функция) не имеет явного аналитического описания, но допускает алгоритмическое представление и требует значительных затрат на расчет. Кроме того, в обратных задачах химической кинетики целевая функция может быть существенно многоэкстремальной, т.е. иметь много локальных экстремумов наряду с глобальным. 
%Часть ННГУ
Численные методы решения подобных многоэкстремальных задач (методы глобальной оптимизации) существенно отличаются от методов локального поиска (see, e.g., \cite{Sergeyev2017,PaulaviciusZilinskas2014}). Как правило, методы локальной оптимизации не могут покинуть области притяжения локального экстремума и не находят глобальный оптимум. А использование параметров модели, соответствующих найденному локальному решению, может оказаться недостаточным, поскольку глобальное решение может дать существенный выигрыш по сравнению с локальным.

Разнообразие возникающих задач глобальной оптимизации влечет за собой разнообразие подходов к их решению.
Методы решения задач глобальной оптмизации можно условно разделить на два класса: метаэвристические и детерминированные. Метаэвристические алгоритмы, как правило, основаны на имитации процессов, протекающих в живой природе.
Примерами метаэвристических алгоритмов являются simulated annealing, evolution and genetic algorithms и т.д. (see, e.g., \cite{Battiti2009,Eiben2015}). В силу своей относительной простоты метаэвристические алгоритмы более популярны у исследователей, чем детерминированные.  Однако решение задачи, найденное метаэвристическим алгоритмом, является, вообще говоря, локальным и может распологаться далеко от глобального \cite{Kvasov2018}.

Возможность построения детерминированных методов глобального поиска, отличающихся от перебора по сетке и метаэвристических алгоритмов, связана с наличием и учетом  априорных предположений о свойствах функций задачи. Такие предположения играют ключевую роль при разработке эффективных алгоритмов глобальной оптимизации и служат основным математическим инструментом для получения оценок глобального решения.

Одним из естественных допущений о задаче является предположение об ограниченности относительных изменений значений целевой функции. Такое допущение связано с тем, что отношения приращений функции к соответствующим приращениям ее аргумента обычно не могут превышать некоторого порога, определяемого ограниченной энергией изменений в моделируемой системе. В таком случае, как известно, функции называются липшицевыми, а сама задача -- задачей липшицевой глобальной оптимизации.

В данной статье отражены результаты применения параллельных методов липшицевой оптимизации, разрабатываемых в ННГУ им. Н.И. Лобачевского, к решению обратных задач химической кинетики. 
Основная часть статьи имеет следующую структуру. Описание математической модели исследуемой химической реакции приведено в Section 2. Формальная постановка задачи липшицевой глобальной оптимизации приводятся в Section 3. В Section 4 кратко изложена схема разрабатываемого авторами асинхронного параллельного алгоритма для решения задач указанного класса. Результаты численного решения обратной задачи химической кинетики обсуждаются в Section 5.

@@

Как уже говорилось выше, с формальной точки зрения the inverse problem of chemical kinetics рассматривается нами как Lipschitz global optimization problem. 
В общем виде задача указанного класса может быть математически сформулирована следующим образом:
\begin{gather}
 \varphi^* = \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}}, \label{problemN}\\
 D=\left\{y\in R^N: a_i\leq y_i \leq b_i, \;  1\leq i \leq N\right\} \label{D},
\end{gather}
где $a,\; b$ есть заданные векторы, $a,b\in R^N$, а целевая функция $\varphi(y)$ удовлетворяет the Lipschitz condition
\begin{equation}\label{Lip}
\left|\varphi(y_1)-\varphi(y_2)\right|\leq L\left\|y_1-y_2\right\|,\; y_1,y_2 \in D.
\end{equation}

Функция $\varphi(y)$ предполагается многоэкстремальной и заданной в виде ``black box'' (т.е. в виде некоторой вычислительной процедуры, на вход которой подается вектор параметров, а на выходе наблюдается соответствующее значение функции). Кроме того, подразумевается, что каждое trial (т.е. вычисление значение функции в точке области поиска) является трудоемкой операцией и требует значительных вычислительных ресурсов. Как отмечалось во введении, такая постановка задачи полностью соответствует the inverse problem of chemical kinetics.

The Lipschitz condition (\ref{Lip}) может быть использовано для оценки глобального минимума функции на интервале, а знание константы Липшица позволяет конструировать алгоритмы глобального поиска и доказывать условия их сходимости (see, for example, \cite{Strongin2000}).

Одной из основных трудностей в решении многомерных задач глобальной оптимизации является рост вычислительных затрат при увеличении размерности задачи. Уменьшение числа испытаний функции при сохранении точности решения возможно за счет более полного использования априорных предположений о целевой функции, что приводит к адаптивным последовательным алгоритмам оптимизации.

Такими методами являются, например, non-uniform space covering method \cite{Evtushenko2013} and simplicial partitions method \cite{Zilinskas2010}. Данные подходы были успешно использованы и для построения параллельных оптимизационных методов \cite{Evtushenko2009,Paulavicius2011}. Другим адаптивным подходом к решению многомерной задачи (\ref{problemN}) является ее редукция к одной или нескольким одномерным задачам с последующим применением одномерных алгоритмов. Такая редукция может быть осуществлена, например, при помощи nested optimization scheme \cite{Grishagin2018} или при помощи Peano-Hilbert curves \cite{Barkalov2018}. Последний из указанных подходов и используется в настоящей работе.

Используя непрерывное однозначное отображение (Peano-Hilbert curve) $y(x)$ отрезка $[0,1]$ вещественной оси на гиперинтервал $D$ из (\ref{D}) можно свести многомерную задачу (\ref{problemN}) к одномерной задаче
\[
\varphi(y^\ast)=\varphi(y(x^\ast))=\min{\left\{\varphi(y(x)): x\in[0,1]\right\}},
\]
where the function $\varphi(y(x))$ will satisfy a uniform H{\"o}lder condition
\[
\left|\varphi(y(x_1))-\varphi(y(x_2))\right|\leq H\left|x_1-x_2\right|^{1/N}
\]
with the H{\"o}lder constant $H$ linked to the Lipschitz constant $L$ by the relation $ H=2 L \sqrt{N+3}$.
Вопросы численного построения различных аппроксимаций Peano-Hilbert curve рассмотрены в \cite{Strongin2000,Sergeyev2013}.

Таким образом, поисковое испытание в некотрой точке $x'\in[0,1]$ будет включать в себя сначала the construction of the image $y'=y(x')$, и лишь затем вычисление значения функции $ z' = \varphi(y')$.

@@

В предлагаемом подходе схема распараллеливания соответствует по принципу ``master/worker''. В процессе-мастере выполняется алгоритм глобального поиска, в котором проводится накопление поисковой информации, оценка на ее основе константы Липшица для целевой функции, определение точек новых испытаний и распределение их по процессам-рабочим. Процессы-рабочие получают от мастера точки, проводят в них новые испытания и отсылают мастеру результаты испытаний.

Будем считать, что на каждой итерации процесс-мастер вычисляет одну точку нового испытания и передает ее для проведения испытания в процесс-рабочий. При этом проведение испытания рабочим является значительно более трудоемкой операцией, чем выбор новой точки мастером, что исключает простой процессов-рабочих. 
В данном случае (в отличие от синхронных параллельных алгоритмов) общее число выполненных испытаний каждым процессом-рабочим будет зависеть от трудоемкости проведения конкретного испытания и не может быть оценено заранее.

При описании параллельного алгоритма будем предполагать, что в нашем распоряжении имеется $p+1$ вычислительный процесс: один мастер и $p$ рабочих.

В начале поиска процесс-мастер (будем считать, что это процесс No 0) инициирует параллельное проведение $p$ испытаний в $p$ различных точках области поиска, две из которых являются граничными, а остальные -- внутренними, т.е. в точках $\{y(x^1), y(x^2), ...,y(x^p)\}$, где $x^1 = 0$, $x^p = 1$, $x^i\in(0,1), i=2,..., p-1$.

Предположим теперь, что выполнено $k$ испытаний (в частности, $k$ может быть равно 0), и процессы-рабочие проводят испытания в точках $\{y(x^{k+1}), y(x^{k+2}), ...,y(x^{k+p})\}$.

Каждый процесс-рабочий, завершивший свое испытание в некоторой точке (не ограничивая общности, будем считать, что это точка $y(x^{k+1})$, соответствующая процессу No 1), пересылает процессу-мастеру результаты испытания. Мастер, в свою очередь, выбирает для него точку нового испытания $x^{k+p+1}$ в соответствии с правилами, описанными ниже.
Отметим, что в данном случае мы будем иметь set of preimages of the trial points
\[
I_k = \left\{ x^{k+1},x^{k+2},...,x^{k+p} \right\},
\]
в которых испытания уже начались, но еще не завершены.

Step 1. Перенумеровать в порядке возрастания (нижним идексом) set of preimages of the trial points 
\[
X_k = \left\{x^1, x^2,...,x^{k+p} \right\},
\]
которое содержит все preimages, в которых либо проведены, либо проводятся испытания, т.е.
\[
0=x_1<x_2<...<x_{x+p}=1.
\]
Step 2. Вычислить значения 
\[
M_1=\max \left\{ \frac{ \left|z_i - z_{i-1} \right|}{(x_i-x_{i-1})^{1/N}} : x_{i-1} \notin I_k, x_i \notin I_k, 2\leq i\leq k+p \right\},
\]
\[
M_2=\max \left\{ \frac{ \left|z_{i+1} - z_{i-1} \right|}{(x_{i+1}-x_{i-1})^{1/N}} : x_i \in I_k, 2\leq i < k+p \right\},
\]
\[
M=\max\{M_1,M_2\},
\]
где $z_i=\varphi(y(x_i))$, if $x_i \notin I_k, \; 1\leq i \leq k+p$. Значения $z_i$ в точках $x_i \in I_k$ являются неопределенными, т.к. испытания в точках $x_i \in I_k$ еще не завершены. Если значение $M$ получилось равным 0, то положить $M=1$.

Step 3. Каждому интервалу $(x_{i-1},x_i), \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p$, поставить в соответствие число $R(i)$, которое называется характеристикой интервала и вычисляется по формуле
\[
R(i)=rM\Delta_i+\frac{(z_i-z_{i-1})^2}{rM\Delta_i}-2(z_i+z_{i-1}),
\]
где $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$, and $r>1$ is the reliability parameter of the method.

Step 4. Выбрать интервал $[x_{t-1},x_t]$, которому соответствует максимальная характеристика, т.е.
\[
R(t) = \max \left\{ R(i): \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p \right\}.
\]

Step 5. Определить точку нового испытания $y^{k+p+1}=y(x^{k+p+1})$, прообраз $x^{k+p+1} \in (x_{t-1},x_t)$ которой в соответствии с формулой
\[
x^{k+p+1} = \frac{x_{t}+x_{t-1}}{2} - \mathrm{sign}(z_{t}-z_{t-1})\frac{1}{2r}\left[\frac{\left|z_{t}-z_{t-1}\right|}{M}\right]^N.
\]

Сразу после вычисления точки очередного испытания процесс-мастер добавляет ее в множество $I_k$ и пересылает процессу-рабочему, который инициирует проведение испытания в ней.

Процесс-мастер завершает работу алгоритма, если выполняется одно из двух условий: $\Delta_{t}<\epsilon$, $k+p>K_{max}$.
Вещественное число $\epsilon>0$ и целое число $K_{max}>0$ являются параметрами алгоритма и соответствуют точности поиска решения и максимальному числу испытаний.

Описанный параллельный асинхнонный алгоритм основан на последовательном information global search algorithm. Теоретическое обоснование сходимости алгоритма приведено в \cite{Strongin2000}. Здесь же приведены схемы синхронного распараллеливания, использовавшиеся ранее при решении ряда прикладных задач \cite{Kalyulin2017,Modorskii2016}.
Новизна данной работы заключается в практической реализации и применении асинхронной схемы распалаллеливания, обладающей большей эффективностью при решении задач с разной трудоемкостью проведения испытаний в разных точках области поиска. Это подтверждается результатами экспериментов, описанными в следующем разделе.

>>>>

%Часть УГНТУ
Currently, there is a tendency to improve the environmental characteristics of automobile fuel while maintaining a high octane number. Sulfuric acid alkylation of isobutane with olefins makes it possible to obtain a high-octane component of gasoline with a minimum content of aromatic hydrocarbons. The alkylate, which is produced by the alkylation of isobutane with C3-C5 olefins in the presence of strong acid, has the advantages of high octane number, low vapor pressure, and zero content of olefins and aromatics that allow it to be a desirable blending component for high-quality gasoline. Alkylates will continue to act as a desirable blending component for high-quality gasoline as the quality of gasoline continues to increase [ссылка на китайскую статью]. Therefore, it is a significant process of the modern refinery [ссылка на catalyst]. To optimize the chemical process in industry, it is necessary to first develop its model, which in this case means building a mathematical model of the chemical process and then its kinetic model, that is, numerically calculate the kinetic constants of the reaction.

%Стыковочный абзац
As a rule, it is impossible to find out the kinetic constants of the reactions analytically. Therefore, there is a need in the development and application of numerical methods for finding the kinetic constants (see, e.g., []). In this case, the quality criteria of the solution found (objective function) haven't an explicitt analytical description but allows an algorithmic representation and requires considerable computaton resources. Moreover, in the inverse problems of chemical kinetics the objective function can be multiextremal essentially, i.e. can have many local extrema along with the global one.

%Часть ННГУ
The numerical methods for solving such multiextremal problems (global optimization methods) differ significantly from the local search ones (see, e.g., \cite{Sergeyev2017,PaulaviciusZilinskas2014}). As a rule, the local optimization methods cannot escape a local extremum attractione region and don't find the global optimum. At the same time, the use of the model parameters corresponding to the local solution found may appear to be insufficient since the global solution may provide a considerable advantage as compared to the local ones.

The diversity of the global optimization problems arising entails various approaches to solving these ones. The methods of solving the global optimization problems can be divided into two classes: the metaheuristic methods and the deterministic ones. The metaheuristic algorithms, as a rule, are based on the simulation of the processes going in the nature. Some examples of the metaheuristic algorithms are simulated annealing, evolution and genetic algorithms, etc. (see e.g. \cite{Battiti2009,Eiben2015}). Because of relative simplicity, the metaheuristic algorithms are more popular among the researchers than the deterministic methods. 
However, a problem solution found by a metaheuristic algorithm is, generally speaking, a local one and may be far from the global solution \cite{Kvasov2018}.

The possibility to construct deterministic global search methods different from the grid search and from the metaheuristic methods is related to the availability of and taking into account some {\it a priori} assumption on the properties of the problem functions. Such assumptions play a key role in the development of efficient global optimization algorithms and serve as main mathematical tools for estimating the global solutions.

An assumption on limited relative variations of the objective function values is one of natural assumptions of a problem. Such an assumption is related to the ratio of the function increment to respective increment of its argument, which is usually limited by some threshold defined by a limited energy of variations in the simulated system. In this case the functions are known as Lipschitz ones and the problem itself is called the Lipschitz global optimization problem.

This paper presents the results of application of parallel Lipschitz optimization methods for solving the inverse problems of chemical kinetics. The main part of the paper has the following structure. The description of the mathematical model of the investigated chemical reaction is presented in Section 2. The formal statement of the Lipschitz global optimization problems and the asynchronous parallel algorithm for solving these ones are described in Section 3. The results of the numerical solving the inverse problem of chemical kinetics are discussed in Section 5.

@@

As it has been already mentioned above, from the formal point of view, we consider the inverse problem of chemical kinetics as Lipschitz global optimization problem. In the general form, a problem of the class specified above can be formulated mathematically as 
follows:
\begin{gather}
 \varphi^* = \varphi(y^\ast)=\min{\left\{\varphi(y):y\in D\right\}}, \label{problemN}\\
 D=\left\{y\in R^N: a_i\leq y_i \leq b_i, \;  1\leq i \leq N\right\} \label{D},
\end{gather}
where $a,b$ are given vectors, $a,b\in R^N$, and the objective function $\varphi(y)$ satisfies the Lipschitz condition
\begin{equation}\label{Lip}
\left|\varphi(y_1)-\varphi(y_2)\right|\leq L\left\|y_1-y_2\right\|,\; y_1,y_2 \in D.
\end{equation}

The function $\varphi(y)$ is assumed to be multiextremal and defined in the form of ``black box'' (i.e. in the form of some computing procedure, the input of which the  vector of parameters is supplied into, and the corresponding function value is taken from the output). Moreover, each \textit{trial} (i.e. the computing of the function value at a point of the search domain) is assumed to be time-consuming operation. 
As it has been noted in Introduction, such problem statement corresponds to the inverse problem of chemical kinetics completely.

The Lipschitz condition (\ref{Lip}) can be utilized to estimate the global minimum of a function within an interval, and knowing the Lipschitz constant allows constructing the global search algorithms and proving the convergence conditions for these ones (see, for example, \cite{Strongin2000}).

The growth of the computation costs with increasing problem dimensionality is one of the main difficulties in solving the multidimensional global optimization problems. The decreasing of the number of trials at preserving the solution accuracy is possible by a complete utilization of some {\it a priori} assumptions on the objective function that leads to adaptive sequential optimization algorithms.

For example, non-uniform space covering method \cite{Evtushenko2013} and simplicial partitions method \cite{Zilinskas2010} are such methods. These approaches were applied successfully for the development of the parallel optimization methods as well \cite{Evtushenko2009,Paulavicius2011}. 
Another adaptive approach to solving a multidimensional problem (\ref{problemN}) is its reduction to a single one-dimensional problem or to several ones followed by application of the one-dimensional algorithms. 
Such a reduction can be made, for example, using nested optimization scheme \cite{Grishagin2018} or Peano-Hilbert curves \cite{Barkalov2018}. 
The latter approach was used in the present work.

Using the continuous unambiguous mapping (Peano-Hilbert curve) $y(x)$ of the interval $[0,1]$ of the real axis on the hypercube $D$ from (\ref{D}), one can reduce a multidimensional problem (\ref{problemN}) to a one-dimensional problem
\[
\varphi(y^\ast)=\varphi(y(x^\ast))=\min{\left\{\varphi(y(x)): x\in[0,1]\right\}},
\]
where the function $\varphi(y(x))$ will satisfy a uniform H{\"o}lder condition
\[
\left|\varphi(y(x_1))-\varphi(y(x_2))\right|\leq H\left|x_1-x_2\right|^{1/N}
\]
with the H{\"o}lder constant $H$ linked to the Lipschitz constant $L$ by the relation $ H=2 L \sqrt{N+3}$. 
The issues of the numerical construction of various approximations of the Peano-Hilbert curve were considered in \cite{Strongin2000,Sergeyev2013}.

So far, a search trial at some point $x'\in[0,1]$ will include first the construction of the image $y'=y(x')$ and then the computing the value of the function $z'=\varphi(y')$.

@@

In the approach proposed, the parallelization scheme corresponds to the ``master/worker'' principle.
In the master process the global search algorithm is executed, which accumulates the search information, evaluates the Lipschitz constant for the objective function on its base, determines the new trial points and distributes these ones among the worker processes.

The worker processes receive the trial points from the master process, perform the new trials at these points, and send the trial results to the master process.

Let us assume the master process to compute one point of the next trial at every iteration and to send it to a worker process for executing the trial. 
At the same time, the execution of the trial by the worker process is much more computational-costly operation than the choice of a new trial point by the master that excludes idle worker processes. 
In this case (unlike the synchronous parallel algorithms), the total number of trials executed by each worker process will depend on the computation costs of executing a particular trial and cannot be estimated in advance.

In the description of the parallel algorithm, let us assume $p+1$ computing processes to be at our disposal: one master process and $p$ worker ones.

At the beginning of the search, the master process (let us assume it to be Process No 0) initiates the parallel execution of $p$ trials at $p$ different points of the search domain. 
Two of these points are the boundary ones while the rest are the internal ones, i.e. at the points $\{y(x^1), y(x^2), ...,y(x^p)\}$ where 
$x^1 = 0$, $x^p = 1$, $x^i\in(0,1), i=2,..., p-1$.

Now let us assume $k\geq 0$ trials (in particular, $k$ may be equal to 0) to be completed, and the worker processes are performing the trials at the points $\{y(x^{k+1}), y(x^{k+2}), ...,y(x^{k+p})\}$.

Each worker process having completed its trial at some point (without any loss of generality, let us assume this point to be $y(x^{k+1})$ corresponding to process No 1), sends the trial result to the master process. 
In turn, the master process selects a new trial point $x^{k+p+1}$ for the worker process according to the rules described below.
Note that in this case we will have a set of preimages of the trial points
\[
I_k = \left\{ x^{k+1},x^{k+2},...,x^{k+p} \right\},
\]
which the trials have been already started at but haven't been completed yet.

Step 1. Renumber set of preimages of the trial points 
\[
X_k = \left\{x^1, x^2,...,x^{k+p} \right\},
\]
containing all the preimages, which the trials either have been completed or are being performed at in the increasing order (by the lower index), so that
\[
0=x_1<x_2<...<x_{x+p}=1.
\]
Step 2. Compute the values 
\begin{gather*}
M_1=\max \left\{ \frac{ \left|z_i - z_{i-1} \right|}{(x_i-x_{i-1})^{1/N}} : x_{i-1} \notin I_k, x_i \notin I_k, 2\leq i\leq k+p \right\},\\
M_2=\max \left\{ \frac{ \left|z_{i+1} - z_{i-1} \right|}{(x_{i+1}-x_{i-1})^{1/N}} : x_i \in I_k, 2\leq i < k+p \right\},\\
M=\max\{M_1,M_2\},
\end{gather*}
where $z_i=\varphi(y(x_i))$ if $x_i \notin I_k, \; 1\leq i \leq k+p$. The values $z_i$ at the points $x_i \in I_k$ are undefined since the trials at the points $x_i \in I_k$ haven't been completed yet. If the value of $M$ equals 0, then set $M=1$.

Step 3. Juxtapose each interval $(x_{i-1},x_i), \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p$ to a quantity $R(i)$, which is called a characteristic of the interval and is computed according to the formula
\[
R(i)=rM\Delta_i+\frac{(z_i-z_{i-1})^2}{rM\Delta_i}-2(z_i+z_{i-1}),
\]
where $\Delta_i=\left(x_i-x_{i-1}\right)^{1/N}$ and $r>1$ is the reliability parameter of the method.

Step 4. Select the interval $[x_{t-1},x_t]$, which the maximum characteristic corresponds to, i.e.
\[
R(t) = \max \left\{ R(i): \; x_{i-1} \notin I_k, x_i \notin I_k, \; 2\leq i\leq k+p \right\}.
\]

Step 5. Define the new trial point $y^{k+p+1}=y(x^{k+p+1})$, the preimage of which is $x^{k+p+1} \in (x_{t-1},x_t)$ according to the formula
\[
x^{k+p+1} = \frac{x_{t}+x_{t-1}}{2} - \mathrm{sign}(z_{t}-z_{t-1})\frac{1}{2r}\left[\frac{\left|z_{t}-z_{t-1}\right|}{M}\right]^N.
\]

Upon computing the next trial point, the master process adds it to the set $I_k$ and sends it to the worker process, which initiates the new trial at this point.

The master process terminates the algorithm if one of two conditions is satisfied: $\Delta_{t}<\epsilon$ or $k+p>K_{max}$.
The real number $\epsilon>0$ and the integer number $K_{max}>0$ are the parameters of the algorithm and correspond to the solution search precision and to the maximum number of trials, respectively.

The parallel asynchronous algorithm described above is based on the sequential information global search algorithm. The theoretical substantiation of the algorithm convergence is given in \cite{Strongin2000}. Also, the synchronous parallelization schemes used earlier in solving a number of applied problems \cite{Kalyulin2017,Modorskii2016} are presented here.
The novelty of the present work consists in a practical implementation and application of the asynchronous parallelization scheme featured by a higher efficiency in solving the problems with different computation costs of performing the trials at different points of the search domain. 
It was confirmed by the results of experiments described in the next section.
